---
title: "TEDS_Nebraska"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r, include= F}
#Clean Environment
rm(list = ls())
```

 
```{r, include=FALSE}
#Load Packages
library(tidyverse)
library(gtsummary)
library(Hmisc)
library(gt)
```

```{r}
#Load Data. Can be downloaded from https://www.datafiles.samhsa.gov/sites/default/files/field-uploads-protected/studies/TEDS-A-2017/TEDS-A-2017-datasets/TEDS-A-2017-DS0001/TEDS-A-2017-DS0001-bundles-with-study-info/TEDS-A-2017-DS0001-bndl-data-r.zip. You will need to save it to your local file and update the file argument.
load(file = "/Users/ryanherrschaft/Documents/TEDS Code/Data/tedsa_puf_2017.RData")
```


```{r import_NE_admissions}
Nebraska_admissions_raw <- tedsa_puf_2017 %>% 
  #This code filters so that only admissions data from Nebraska are included.
  filter(STFIPS == 31) %>% 
  #In the original dataset, there are 62 variables, but many of these variables are not of interest to us. I have included the following variables:
       # Participant ID (CASEID),
        #Micro/Metropolitian Area (CBSA2010),
        #Service Setting (SERVICES),
        #Referral source (PSOURCE)
        #Participant Gender (GENDER)
        #Substances Used at time of admission (SUB1:3)
        #Usual Route of Substance Administration (ROUTE1:3),
        #Substance Usage dummy coded (All variables ending in FLG), 
        #and an indicator on combined alcohol/other drug use (ALCDRUG).
  select(CASEID, CBSA2010, SERVICES, PSOURCE, GENDER, starts_with("SUB"), starts_with("ROUTE"), ends_with("FLG"), ALCDRUG)
```
 
```{r recode_and_clean_admissions, include = F}
## Recoding 
#In order to make the data easy to interpret and manipulate, I converted the datatype of categorical variables to the "factor" datatype and provided more intuitive labels.
NE_admissions_cleaned <- Nebraska_admissions_raw %>% 
  mutate(
    CASEID = factor(CASEID),
#Recode CBSA from the 13 5 digit codes within Nebraska to three categories: Outside CBSA, Micropolitan CBSA, and Metropolitan CBSA. This recoding was done by labeling all the CBSA 5-digit codes with the CBSA category pulled from this chart: https://www.uspto.gov/web/offices/ac/ido/oeip/taf/cls_cbsa/cbsa_countyassoc.htm. 
    CBSA2010_trichot = recode(CBSA2010,
      "-9" = "Outside CBSA",
      "13100" = "Micropolitan CBSA",
      "18100" = "Micropolitan CBSA",
      "23340" = "Micropolitan CBSA",
      "24260" = "Micropolitan CBSA",
      "25580" = "Micropolitan CBSA",
      "28260" = "Micropolitan CBSA", 
      "30420" = "Micropolitan CBSA", 
      "35740" = "Micropolitan CBSA", 
      "35820" = "Micropolitan CBSA", 
      "42420" = "Micropolitan CBSA", 
      "30700" = "Metropolitan CBSA",
      "36540" = "Metropolitan CBSA",
      "43580" = "Metropolitan CBSA",
      ), .after = CBSA2010,

#Recoding the Service, Gender, Route, and Substance variables to have more informative labels.
    SERVICES = factor(SERVICES) %>% recode(
          "1" = "Detox, 24-hour, hospital inpatient",
          "2" = "Detox, 24-hour, free-standing residential",
          "3" = "Rehab/residential, hospital (non-detox)",
          "4" = "Rehab/residential, short term (30 days or fewer)",
          "5" = "Rehab/residential, long term (30 days or more)",
          "6" = "Ambulatory, intensive outpatient",
          "7" = "Ambulatory, non-intensive outpatient",
           "8" = "Ambulatory, detox"
                             ),

    GENDER = factor(GENDER) %>%  recode(
                            "1" = "Male",
                            "2" = "Female",
                            "-9" = "NA"
                            ),
    ROUTE1 = factor(ROUTE1) %>% recode(
                            "1" = "Oral",
                            "2" = "Smoking",
                            "3" = "Inhalation",
                            "4" = "Injection (IV or intramuscular)",
                            "5" = "Other",
                            "-9" = "NA"
                            ),
    ROUTE2 = factor(ROUTE2) %>% recode(
                            "1" = "Oral",
                            "2" = "Smoking",
                            "3" = "Inhalation",
                            "4" = "Injection (IV or intramuscular)",
                            "5" = "Other",
                            "-9" = "NA"
                            ),
    ROUTE3 = factor(ROUTE3) %>% recode(
                            "1" = "Oral",
                            "2" = "Smoking",
                            "3" = "Inhalation",
                            "4" = "Injection (IV or intramuscular)",
                            "5" = "Other",
                            "-9" = "NA"
                            ),
    SUB1 = factor(SUB1) %>% recode(
                            "1" = "None",
                            "2" = "Alcohol",
                            "3" = "Cocaine/crack",
                            "4" = "Marijuana",
                            "5" = "Heroin",
                            "6" = "Non-prescription methadone",
                            "7" = "Synthetic opioids",
                            "8" = "PCP",
                            "9" = "Other hallucinogens",
                            "10" = "Methamphetamine",
                            "11" = "Other amphetamines",
                            "12" = "Other stimulants",
                            "13" = "Benzodiazepine", 
                            "14" = "Other non-benzodiazepine tranquilizers",
                            "15" = "Barbiturates",
                            "16" = "Other non-barbiturate sedatives or hypnotics",
                            "17" = "Inhalants",
                            "18" = "Over-the-counter medications",
                            "19" = "Other",
                            "-9" = "Missing/unknown/not collected/invalid"
                            ),
        SUB2 = factor(SUB2) %>% recode(
                             "1" = "None",
                            "2" = "Alcohol",
                            "3" = "Cocaine/crack",
                            "4" = "Marijuana",
                            "5" = "Heroin",
                            "6" = "Non-prescription methadone",
                            "7" = "Synthetic opioids",
                            "8" = "PCP",
                            "9" = "Other hallucinogens",
                            "10" = "Methamphetamine",
                            "11" = "Other amphetamines",
                            "12" = "Other stimulants",
                            "13" = "Benzodiazepine", 
                            "14" = "Other non-benzodiazepine tranquilizers",
                            "15" = "Barbiturates",
                            "16" = "Other non-barbiturate sedatives or hypnotics",
                            "17" = "Inhalants",
                            "18" = "Over-the-counter medications",
                            "19" = "Other",
                            "-9" = "Missing/unknown/not collected/invalid"
                            ),
        SUB3 = factor(SUB3) %>% recode(
                            "1" = "None",
                            "2" = "Alcohol",
                            "3" = "Cocaine/crack",
                            "4" = "Marijuana",
                            "5" = "Heroin",
                            "6" = "Non-prescription methadone",
                            "7" = "Synthetic opioids",
                            "8" = "PCP",
                            "9" = "Other hallucinogens",
                            "10" = "Methamphetamine",
                            "11" = "Other amphetamines",
                            "12" = "Other stimulants",
                            "13" = "Benzodiazepine", 
                            "14" = "Other non-benzodiazepine tranquilizers",
                            "15" = "Barbiturates",
                            "16" = "Other non-barbiturate sedatives or hypnotics",
                            "17" = "Inhalants",
                            "18" = "Over-the-counter medications",
                            "19" = "Other",
                            "-9" = "Missing/unknown/not collected/invalid"),
#For our purposes, we used a more expansive coding of Other than the original dataset. Our designation of Other included Other non-barbiturate sedatives/hypnotics (SEDHPFLG), Inhalants (INHFLG), Over the counter medications (OTCFLG), Other drugs (OTHERFLG), Barbiturates (BARBFLG), Other non-benzodiazepine tranquilizers (TRNQFLG), Other stimulants (STIMFLG), Other hallucinogens (HALLFLG), PCP (PCPFLG), Non-prescription methadone (METHFLG), and Other amphetamines (AMPHFLG).
        other_drug_use_FLG = (SEDHPFLG + INHFLG + OTCFLG + OTHERFLG + BARBFLG + TRNQFLG + STIMFLG +
                       HALLFLG + PCPFLG + METHFLG + AMPHFLG) %>%  recode(`0` = 0, .default = 1),
#Coding a None dummy variable
        none_FLG = recode(ALCDRUG, 
                          "0" = 1, #Recode No Drug use to a 1 indicator
                          "1" = 0, #Else = 0
                          "2" = 0,
                          "3" = 0),

        PSOURCE = recode(PSOURCE,
                         "7" = "Court/criminal justice referral/DUI/DWI",
                         "1" = "Individual (includes self-referral)",
                         "2" = "Alcohol/drug use care provider",
                         "3" = "Other health care provider",
                         "4" = "School",
                         "5" = "Employer/EAP",
                         "6" = "Other Community Referral",
                         "-9" = "NA"
                         )
) %>% 
  #Reorder Columns to place new FLG variables at the end.
  select(CASEID:CBSA2010_trichot, SERVICES:ALCDRUG, other_drug_use_FLG, none_FLG )

```


```{r}
#Here you can look at the dataset
head(NE_admissions_cleaned)

```


```{r convert_to_long, include = F}
## Data Tidying 

#In order to conduct some visualizations and manipulations in the tidyverse, data must be in the tidy format. The  tidy format means that each observation takes up one row. I call this long instead of tidy, becasue one admission stretches across many observations.

NE_drugs_long <- NE_admissions_cleaned %>% 
  #Remove all variables that will not be used for ggplot visualziations. THis includes the drug variables that have been recoded to the "other_drug_use_FLG".
  select(-starts_with("SUB"), -starts_with("ROUTE"), -METHFLG, -PCPFLG:-HALLFLG, -AMPHFLG:-STIMFLG, -TRNQFLG:-OTHERFLG, -ALCDRUG) %>%  
#Pivot longer takes all of the substance use flags (ends_with("FLG")) and creates unique rows for each. I use pivot_longer here for compatibility with the ggplot2 data visualzation package. In the resulting dataset, each row will specify whether or not a patient mentioned a specific substance at admission. GGPlot requires information to be in this form (often called the tidy format) to take advantage of its full capabilities. For more https://tidyr.tidyverse.org/articles/pivot.html.
  pivot_longer(cols = ends_with("FLG")) %>%
  rename(SUBSTANCE = name, USAGE = value) %>%
  mutate(
#Recode Substances to meaningful names
    SUBSTANCE = factor(SUBSTANCE) %>% 
    recode(
      "MTHAMFLG" = "Methamphetamine",
      "MARFLG" = "Marijuana",
      "HERFLG" = "Heroin",
      "COKEFLG" = "Cocaine/crack",
      "ALCFLG" = "Alcohol",
      "BENZFLG" = "Benzodiazepine",
      "OPSYNFLG" = "Synthetic opioids",
      "other_drug_use_FLG" = "Other",
      "none_FLG" = "None"
          )
        ) 
```

```{r}
#Take a look at the new data frame. Notice that each row references a unique substance-usage combination.
head(NE_drugs_long)
```


```{r read_in_census_data, include = F}
## Census data ##

# To compute population adjusted substance use reporting rates across CBSA_2010 categories, I needed to import Census data that captured the estimated population by CBSA region. This data was not directly available. To capture it, this section of code imports NE county population estimates from the Census and specifies whether a county is metro/micro/neither. Later aggregates use those tables to calculate the population statewide and by CBSA category.

#read in data on the counties and whether they are metro or micro. I developed this csv myself after looking at a map of the Nebraskan CBSA regions: https://www2.census.gov/geo/maps/metroarea/stcbsa_pg/Feb2013/cbsa2013_NE.pdf. Dark green counties were labeled Metropolitan, light green micropolitan. For access to this file, contact Ryan Herrschaft (rherrschaft2@unl.edu). You will need to update the file path to reflect the documents location on your computer.
CBSA_regions <- read_csv(file ="/Users/ryanherrschaft/Documents/TEDS Code/Data/CBSA for r.csv") 

#read in census data on the population of the counties. This data is available at https://www2.census.gov/programs-surveys/popest/datasets/2010-2017/counties/asrh/cc-est2017-alldata-31.csv. You may need to ask me for this datafile, save it to your local computer, and then change the file argument to reflect where you downloaded this on your computer.
census_pop_data_raw <- read_csv("/Users/ryanherrschaft/Documents/TEDS Code/Data/cc-est2017-alldata-31.csv")

#Fiter census data for the population 10 and over, find the population by county.
census_pop_10_and_over <- census_pop_data_raw %>% 
  filter(AGEGRP >=  3 & YEAR == 10) %>% #The age filter captures everyone over 10, the year variable specifies the year 2017.
  select(CTYNAME, AGEGRP, TOT_POP) %>% 
#These next two commands in tandem provide the over 10 population of each county in the year 2017.
  group_by(CTYNAME) %>% 
  summarise(sum_population_of_county_10_and_older = sum(TOT_POP))

 
#Fiter census data for the population 15 and over, find the population by county. 
census_pop_15_and_over <- census_pop_data_raw %>%
  filter(AGEGRP >=  4 & YEAR == 10) %>% 
   select(CTYNAME, AGEGRP, TOT_POP) %>% 
  group_by(CTYNAME) %>%
  summarise(sum_population_of_county_15_and_older = sum(TOT_POP) ) 
  
#Join the two age-based population counties together, prepare County names for join by making their name in the same format as the CBSA_regions df.
county_population_table_intermediate <- left_join(census_pop_10_and_over, census_pop_15_and_over, by = "CTYNAME") %>% 
  arrange(desc(sum_population_of_county_15_and_older)) %>%
  mutate(CTYNAME = str_remove_all(CTYNAME, pattern = " County"))

#Join the population counts of each county with the df specifying the CBSA region of that county.
county_population_table <- county_population_table_intermediate %>% 
  left_join(CBSA_regions, by = "CTYNAME") %>% 
  #Replace counties not in the CBSA regions dataset with the Outside CBSA label instead of NA.
  mutate(CBSA = replace_na(CBSA, "Outside CBSA") )

#Aggregate up from the county level population sums to the CBSA type population sums.
CBSA_population_table <- county_population_table %>%
  group_by(CBSA) %>%
  summarise(pop_of_CBSA_10_and_older = sum(sum_population_of_county_10_and_older),     
            pop_of_CBSA_15_and_older = sum(sum_population_of_county_15_and_older)) %>% 
  mutate(CBSA = recode(CBSA, 
         "Metro" = "Metropolitan CBSA",
         "Micro" = "Micropolitan CBSA")) #Recode names so they are compatible for later joins by CBSA
```

```{r}
#Useful Denominators

#Computing population of Nebraska, statewide. This will be the denominator when I later compute the rate of admissions per 100k of the state population. 
Nebraska_pop_over_10 <- sum(census_pop_10_and_over$sum_population_of_county_10_and_older)
Nebraska_pop_over_15 <- sum(census_pop_15_and_over$sum_population_of_county_15_and_older)

#This will be used in many percent codes. It refers to the number of unique admissions
num_unique_admissions <- length(NE_admissions_cleaned$CASEID)
```

 
 
```{r summary_table_statewide}
### Substance Use ###
#Here is a summary table showing the number of people reporting the use of each type of drug at time of admission. 
substance_use_summary_table <- NE_drugs_long %>% 
  count(SUBSTANCE, USAGE) %>% 
  filter(USAGE == 1) %>%
  rename(n_reporting_sub_use = n) %>% 
  mutate(
  #Calculate percent of admittees using each substance
        percent = ((n_reporting_sub_use / num_unique_admissions) * 100) %>% round(2),
 #Calculate rate of admittees using each substance per 100k people in state over 10
        per_100k_over_10 = ( (n_reporting_sub_use / Nebraska_pop_over_10) * 100000 ) %>%  round(2)
        ) 

```


```{r}
relevel_order <- c("None", "Other", "Benzodiazepine","Heroin", "Cocaine/crack","Synthetic opioids", "Methamphetamine", "Marijuana", "Alcohol")
```


### Figure 5: All substances, Statewide, Percent
```{r all_sub_statewide_percent_graph, include = F}
#Lolipop graphs may not be the simplest way to represent the data; the line segments and labels can make them appear quite busy. However, I find that the redundancy of information display (ie., same information conveyed by locaiton of point on x axis, label, and length of line segment) helps make the data easier to interpret. A strictly academic audience may prefer just the point. I can make that change easily by removing the geom_segment and geom_text code.

all_sub_statewide_percent_graph <- ggplot(substance_use_summary_table, aes(x = fct_relevel(SUBSTANCE, relevel_order),
                 y = percent)) +
    geom_point() +
    geom_text(aes(label = paste0(percent, "%")), nudge_y = 6) + 
    geom_segment(aes(x = SUBSTANCE, 
                   xend= SUBSTANCE, 
                   y = 0, 
                   yend = percent) 
                 ) +
    ylim(0,100) +
      labs(
    x = "Substance",
    y = "Percent of Admissions",
    title = "Figure 5: Percent of Admissions Listing Substance at Admission to NE \nTreatment Centers in 2017"
  ) +
    coord_flip() +
    theme_classic() 
```
```{r}
all_sub_statewide_percent_graph
ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_5_all_sub_statewide_percent_graph.png", plot = all_sub_statewide_percent_graph)
```


### Figure 7: All substance, Statewide, per 100k 10+
```{r all_sub_statewide_per_100k_10_plus_graph, include = F}
all_sub_statewide_per_100k_10_plus_graph <- ggplot(substance_use_summary_table, aes(x = fct_relevel(SUBSTANCE, relevel_order),
                 y = per_100k_over_10)) +
    geom_point() +
    geom_text(aes(label = per_100k_over_10), nudge_y = 40) + 
    geom_segment(aes(x = SUBSTANCE, 
                   xend= SUBSTANCE, 
                   y = 0, 
                   yend = per_100k_over_10) 
                 ) +
      labs(
    x = "Substance",
    y = "Admissions per 100,000 people",
    title = "Figure 7: Rate of Admissions Listing Substance at Admission \nto NE Treatment Centers in 2017 per 100,000 Nebraskans",
    caption = "Facilities were limited to listing at most three substances"
  ) +
    coord_flip() +
    theme_classic() 
```
```{r}
all_sub_statewide_per_100k_10_plus_graph

ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_7_all_sub_statewide_per_100k_10_plus_graph.png", plot = all_sub_statewide_per_100k_10_plus_graph)
```


```{r all_sub_use_by_CBSA_table, include = F}

#This counts the number of admissions occuring in each CBSA region. The values stored in `n_admissions_in_stat_area` will be used later to compute the number of admissions (n_admissions_in_stat_area) per 100k in the CBSA region.
admissions_per_CBSA <- NE_admissions_cleaned %>%   
  group_by(CBSA2010_trichot) %>% 
  count() %>%  
  rename(n_admissions_in_stat_area = n) #renames 'n' to a more intuitive name
 

#This table computes similar statistics as the previous table (percent reporting substance use and rate of admissions per 100k), but instead separates these statistics out according to the CBSA region.
substance_use_by_CBSA_summary_table <- 
  NE_drugs_long %>% 
#Counts the number of admissions that occured per substance, per CBSA.
  count(SUBSTANCE, USAGE, CBSA2010_trichot) %>%  
#Rename count to an intutive name
  rename(n_reporting_sub_use = n) %>% 
#We are only interested in visualizing the admissions that used a substance, not the number of admissions that did not use a substance. For this reason, I filter out the values indicating no usage.
  filter(USAGE == 1) %>% 
#Next, I join this table with a dataframe specifying the number of admissions per CBSA region. This variable will serve as the denominator for the percent of admissions per CBSA region reporting using a particular substance.
  left_join(admissions_per_CBSA, by = "CBSA2010_trichot") %>% 
#Computing the percent of admittees per CBSA region that reported using each substacne () 
  mutate(
    percent_of_admittees_in_stat_area_rept_sub_use = (n_reporting_sub_use * 100 / n_admissions_in_stat_area) %>% round(2) ,
  #create labels for the graph that show the number of admissions per CBSA region.  
    CBSA2010_trichot_labels = paste0(CBSA2010_trichot, ", n = ", n_admissions_in_stat_area)
  ) 

substance_use_by_CBSA_summary_table <- substance_use_by_CBSA_summary_table %>%  
#I am joining the last table with a dataframe containing the population of each CBSA region. This will serve as the denominator for computing the rate of admissions for each substance per the 100k population of the region.
  left_join(CBSA_population_table, by = c("CBSA2010_trichot" = "CBSA")) %>% 
  mutate(
    count_per_100k_10_plus = ((n_reporting_sub_use / pop_of_CBSA_10_and_older) * 100000 ) %>% round(2),
    count_per_100k_15_plus = ((n_reporting_sub_use / pop_of_CBSA_15_and_older) * 100000 ) %>% round(2)
        )
```

### Figure 6: All substances, CBSA Areas , percent
```{r all_sub_by_CBSA_percent_graph, include = F}
all_sub_by_CBSA_percent_graph <- ggplot(substance_use_by_CBSA_summary_table, aes(x = fct_relevel(SUBSTANCE, relevel_order), y = percent_of_admittees_in_stat_area_rept_sub_use)) +
    geom_point() +
    geom_text(aes(label = paste0(percent_of_admittees_in_stat_area_rept_sub_use, "%")), nudge_y = 18, size = 3) +
    geom_segment(aes(x = SUBSTANCE, 
                   xend= SUBSTANCE, 
                   y = 0, 
                   yend = percent_of_admittees_in_stat_area_rept_sub_use)
                 ) +
    ylim(0,100) +
      labs(
    x = "Substance",
    y = "Percent of Admissions",
    title = "Figure 6: Percent of Admissions Listing Substance at Admission to NE \nTreatment Centers in 2017, by CBSA Region"
  )+
  facet_wrap(~ CBSA2010_trichot_labels)+
    coord_flip() +
    theme_classic() 
```
```{r}
all_sub_by_CBSA_percent_graph

ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_6_all_sub_by_CBSA_percent_graph.png", plot = all_sub_by_CBSA_percent_graph)
```


## Figure 8: All substances, by CBSA, per 100k 
```{r all_sub_by_CBSA_per_100k_10plus_graph, include = F}
all_sub_by_CBSA_per_100k_10plus_graph <- ggplot(substance_use_by_CBSA_summary_table, aes(x = fct_relevel(SUBSTANCE, relevel_order), y =count_per_100k_10_plus )) +
  geom_point() +
    geom_text(aes(label = count_per_100k_10_plus) , nudge_y = 160 , size = 3) +
    geom_segment(aes(x = SUBSTANCE, 
                   xend= SUBSTANCE, 
                   y = 0, 
                   yend = count_per_100k_10_plus)
                 ) +
      labs(
    x = "Substance",
    y = "Admissions per 100,000 people",
    title = "Figure 8: Rate of Admissions Listing Substance at Admission \nto NE Treatment Centers in 2017 per 100,000 Nebraskans, \nby CBSA Region"
  )+
  facet_wrap(~ CBSA2010_trichot_labels) +
  ylim(0, 1250) +
    coord_flip() +
    theme_classic() +
  theme(panel.spacing.x = unit(.5, "lines")) # Last minute adjustments to get plot to fit in frame. May need to be changed.
```
```{r}
all_sub_by_CBSA_per_100k_10plus_graph

ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_8_all_sub_by_CBSA_per_100k_10plus_graph.png", all_sub_by_CBSA_per_100k_10plus_graph)
```


# End of All Substances. Code Beneath this marker focuses on Primary Substance Only

```{r primary_sub_table, include = F}

#I add less anotations here, becasue this code follows a very similar procedure as the code that produced the tables for all substances. The difference here is that we are only including what the admittee reported as their primary substance. 

primary_substance <- NE_admissions_cleaned %>%  
#Select variables of interest.
  select(CASEID:SUB1, CBSA2010_trichot) %>% 
#Recode to "other" category.
  mutate(SUB1_recoded = recode(SUB1, 
                            "Non-prescription methadone" = "Other",
                            "PCP" = "Other",
                           "Other hallucinogens" = "Other",
                            "Other amphetamines" = "Other",
                            "Other stimulants" = "Other",
                           "Other non-benzodiazepine tranquilizers" = "Other",
                            "Barbiturates" = "Other",
                            "Other non-barbiturate sedatives or hypnotics" = "Other",
                            "Inhalants" = "Other",
                            "Over-the-counter medications" = "Other",
                            "Missing/unknown/not collected/invalid"= "NA")
  ) %>% 
#Count the number of addmitees Listing each substance as their primary substance.
  count(SUB1_recoded) %>% 
  rename(n_reporting_sub_use = n) %>% 
#Create variables displaying: Admissions reporting each substance as their primary substance as a percent of total admissions and a rate of admissions per 100k Nebraskans.
      mutate(
           percent = ((n_reporting_sub_use / num_unique_admissions) * 100 )%>% round(2), 
           per_100k_over_10 = ((n_reporting_sub_use / Nebraska_pop_over_10) * 100000) %>%  round(2),
           per_100k_over_15 = ((n_reporting_sub_use / Nebraska_pop_over_15) * 100000) %>% round(2)
           ) %>% 
  filter(SUB1_recoded != "NA") #Remove NA.

```


### Figure 3: Primary Substance, Statewide per 100k 10+

```{r primary_sub_statewide_per_100k_10plus_graph, include = F }
primary_sub_statewide_per_100k_10plus_graph <- ggplot(primary_substance, aes(x = fct_relevel(SUB1_recoded, relevel_order),
                 y = per_100k_over_10)) +
    geom_point() +
    geom_text(aes(label = per_100k_over_10), size = 3, nudge_y = 25) + 
    geom_segment(aes(x = SUB1_recoded, 
                   xend= SUB1_recoded, 
                   y = 0, 
                   yend = per_100k_over_10) 
                 ) +
      labs(
    x = "Primary Substance",
    y = "Admissions per 100,000 people",
    title = "Figure 3: Rate of Individuals Listing Primary Substance Use at \nAdmission to NE Treatment Centers in 2017 per 100,000 Nebraskans"
  ) +
  ylim(0,550) +
    coord_flip() +
    theme_classic() 
```
```{r}
primary_sub_statewide_per_100k_10plus_graph
ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_3_primary_sub_statewide_per_100k_10plus_graph.png", primary_sub_statewide_per_100k_10plus_graph)
```




### Figure 1: Primary Substance, Statewide percent
```{r primary_sub_statewide_percent_graph, include = F}
primary_sub_statewide_percent_graph <- ggplot(primary_substance, aes(x = fct_relevel(SUB1_recoded, relevel_order ),
                 y = percent)) +
    geom_point() +
    geom_text(aes(label = paste0(percent, "%")), size = 3, nudge_y = 5) + 
    geom_segment(aes(x = SUB1_recoded, 
                   xend= SUB1_recoded, 
                   y = 0, 
                   yend = percent) 
                 ) +
    ylim(0,100) +
      labs(
    x = "Primary Substance",
    y = "Percent of Admissions",
    title = "Figure 1: Percent of Admissions Listing Drug as Primary Substance at \nAdmission to NE Treatment Centers in 2017"
  ) +
    coord_flip() +
    theme_classic() 
```
```{r}
primary_sub_statewide_percent_graph
ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_1_primary_sub_statewide_percent_graph.png", primary_sub_statewide_percent_graph)
```


```{r primary_sub_by_CBSA_table, include = F}
primary_substance_by_CBSA <- NE_admissions_cleaned %>%
#Select the variables of interest.
  select(CASEID:SUB1, CBSA2010_trichot) %>%  
#Recode the substance use variable to create an Other category. We defined this other category based on the drugs that were infrequently reported. Infrequently was determined to be everything that was mentioned less than 1% at admission.
  mutate(SUB1_recoded = recode(SUB1, 
                            "Non-prescription methadone" = "Other",
                            "PCP" = "Other",
                           "Other hallucinogens" = "Other",
                            "Other amphetamines" = "Other",
                            "Other stimulants" = "Other",
                           "Other non-benzodiazepine tranquilizers" = "Other",
                            "Barbiturates" = "Other",
                            "Other non-barbiturate sedatives or hypnotics" = "Other",
                            "Inhalants" = "Other",
                            "Over-the-counter medications" = "Other",
                            "Missing/unknown/not collected/invalid"= "NA")
  ) %>%  
#Count the number of admissions by susbtance adn CBSA region.
  count(SUB1_recoded, CBSA2010_trichot) %>% 
#Rename the generic "n" variable to a more descriptive name.
  rename(n_reporting_sub_use = n) %>%  
#Join the data on the number of admissions per CBSA region. This will allow us to compute the percent of admissions per CBSA that reported this drug as the primary substance at admission.
  left_join(admissions_per_CBSA, by = "CBSA2010_trichot") %>%  
#Join the data on admissions to the data that counts the population of each CBSA. This will enable us to report admissions in terms of population per 100k.
  left_join(CBSA_population_table, by = c("CBSA2010_trichot" = "CBSA")) %>%  
   mutate(
#Compute the number of people reporting a substance as their primary substance and divide by the population of the region in which that admission occured. This gives a ratio of the number of admissions per substance per region to the number of people lving in that region. 
     count_per_100k_10_plus = ((n_reporting_sub_use / pop_of_CBSA_10_and_older) * 100000 ) %>% round(2),
#compute the percent of admissions per CBSA that reported this drug as the primary substance at admission.
    percent_substance_per_CBSA =  ((n_reporting_sub_use / n_admissions_in_stat_area) * 100) %>% round(2),
#add labels for graph.
    CBSA2010_trichot_labels = paste0(CBSA2010_trichot, ", n = ", n_admissions_in_stat_area) 
         ) %>% 
  filter(SUB1_recoded != "NA") #Remove NA
```


### Figure 4: Primary Substance, by CBSA Areas, per 100k 10+
 
```{r primary_sub_by_CBSA_per_100k_10plus_graph}
primary_sub_by_CBSA_per_100k_10plus_graph <- ggplot(primary_substance_by_CBSA, aes(x = fct_relevel(SUB1_recoded, relevel_order), y =count_per_100k_10_plus )) +
  geom_point() +
    geom_text(aes(label = count_per_100k_10_plus) , nudge_y = 130, size = 3) +
    geom_segment(aes(x = SUB1_recoded, 
                   xend= SUB1_recoded, 
                   y = 0, 
                   yend = count_per_100k_10_plus)
                 ) +
      labs(
    x = "Primary Substance",
    y = "Admissions per 100,000 people",
    title = "Figure 4: Rate of Admissions Listing Drug as Primary Substance at \nAdmission to NE Treatment Centers in 2017 per 100,000 Nebraskans,\nby CBSA Region"
  ) +
  ylim(0,1000) +
  facet_wrap(~ CBSA2010_trichot_labels)+
    coord_flip() +
    theme_classic() 
```
```{r}
primary_sub_by_CBSA_per_100k_10plus_graph
ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_4_primary_sub_by_CBSA_per_100k_10plus_graph.png", plot = primary_sub_by_CBSA_per_100k_10plus_graph)
```

## Figure 2: Primary Substance, by CBSA Areas, Percent
```{r primary_sub_by_CBSA_percent_graph, include = F}
primary_sub_by_CBSA_percent_graph <- ggplot(primary_substance_by_CBSA, aes(x = fct_relevel(SUB1_recoded, relevel_order), y = percent_substance_per_CBSA)) +
  geom_point() +
    geom_text(aes(label = paste0(percent_substance_per_CBSA, "%")) , nudge_y = 15, size = 3) +
    geom_segment(aes(x = SUB1_recoded, 
                   xend= SUB1_recoded, 
                   y = 0, 
                   yend = percent_substance_per_CBSA)
                 ) +
      labs(
    x = "Primary Substance",
    y = "Percent of Admissions",
    title = "Figure 2: Percent of Admissions Listing Drug as Primary Substance at \nAdmission to NE Treatment Centers in 2017, by CBSA Region"
  ) +
  ylim(0,100) +
  facet_wrap(~ CBSA2010_trichot_labels)+
    coord_flip() +
    theme_classic()
```
```{r}
primary_sub_by_CBSA_percent_graph
ggsave(filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/Figure_2_primary_sub_by_CBSA_percent_graph.png", plot = primary_sub_by_CBSA_percent_graph)
```



## Table on Polysubstance Use Across CBSA Regions
```{r}
dataset_with_amt_drugs_mentioned <- NE_admissions_cleaned %>% group_by(CASEID) %>%  mutate(num_drugs_mentioned = sum(ALCFLG, COKEFLG, MARFLG, HERFLG, METHFLG, OPSYNFLG, PCPFLG, HALLFLG, MTHAMFLG, AMPHFLG, STIMFLG, BENZFLG, TRNQFLG, BARBFLG, SEDHPFLG, INHFLG, OTCFLG, OTHERFLG)) %>%  ungroup()

label(dataset_with_amt_drugs_mentioned$num_drugs_mentioned) <- "Number of Substances Listed at Admission"

 polysub_use_across_CBSA <- dataset_with_amt_drugs_mentioned %>% 
  select(CBSA2010_trichot, num_drugs_mentioned) %>% 
  tbl_summary(by = CBSA2010_trichot) %>% 
   add_overall() %>%
   modify_header(label = "****") %>% 
     as_gt() %>% 
  tab_header("Table 1: Number of substances listed at admission by CBSA category")


gtsave(data = polysub_use_across_CBSA, filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/polysub_use_across_CBSA.html")
```


Data on Incarceration
```{r}
incarceration_ref_by_subsatance_data <- NE_drugs_long %>% filter(USAGE == 1) %>% 
     mutate(PSOURCE_mod = 
           if_else(PSOURCE == "Court/criminal justice referral/DUI/DWI", "Criminal Justice System", "Other")) 

label(incarceration_ref_by_subsatance_data$PSOURCE_mod) <- "Referral Source"


incarceration_ref_by_subsatance_tbl  <- incarceration_ref_by_subsatance_data %>% 
    select(SUBSTANCE, PSOURCE_mod) %>% 
    tbl_summary(by = SUBSTANCE) %>% 
    as_gt()
    

  gtsave(data = incarceration_ref_by_subsatance_tbl, filename = "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/incarceration_ref_by_subsatance_tbl.html")
```

Looking at the distribution of the Other Category across the CBSA regions.  
```{r}
#In this dataset, I will look only at those drugs in the other category

other_disagregated <- NE_admissions_cleaned %>% 
  select(-starts_with("SUB"), -starts_with("ROUTE"), -CBSA2010:-GENDER, METHFLG, PCPFLG:HALLFLG, AMPHFLG:STIMFLG, TRNQFLG:OTHERFLG, -ALCDRUG, CBSA2010_trichot) %>%  
  pivot_longer(cols = ends_with("FLG")) %>%
  rename(SUBSTANCE = name, USAGE = value) %>%
  mutate(
#Recode Substances to meaningful names
    SUBSTANCE = factor(SUBSTANCE) %>% 
    recode(
      "MTHAMFLG" = "Methamphetamine",
      "MARFLG" = "Marijuana",
      "HERFLG" = "Heroin",
      "COKEFLG" = "Cocaine/crack",
      "ALCFLG" = "Alcohol",
      "BENZFLG" = "Benzodiazepine",
      "OPSYNFLG" = "Synthetic opioids",
      "other_drug_use_FLG" = "Other",
      "none_FLG" = "None"
          )
        )  %>% 
  filter(USAGE == 1)


other_drugs_by_CBSA <- other_disagregated %>% 
  #Filter for only drugs in the other category.
  filter(SUBSTANCE %in% c("SEDHPFLG",  "INHFLG", "OTCFLG" ,"OTHERFLG", "BARBFLG", "TRNQFLG", "STIMFLG", "HALLFLG", "PCPFLG", "METHFLG", "AMPHFLG" )) %>%  
  select(-CASEID, -USAGE) %>% 
  tbl_summary(by = CBSA2010_trichot) %>% 
  add_overall() %>% 
  as_gt()

gtsave(other_drugs_by_CBSA, "/Users/ryanherrschaft/Documents/TEDS Code/Graphs and Figures/other_by_CBSA_tbl.html")

```

### Extra  Information
```{r, include = F}
#Admissions Across Service Settings.
admissions_to_service_sites_by_CBSA <- NE_admissions_cleaned %>% 
  count(SERVICES, CBSA2010_trichot) %>% 
  rename(n_using_service_type = n ) %>% 
  left_join(admissions_per_CBSA, by = "CBSA2010_trichot") %>% 
  mutate(percent_of_admittees_in_stat_area_using_service_type = (n_using_service_type * 100 / n_admissions_in_stat_area) %>% round(2) ) %>% 
  mutate(CBSA2010_trichot_labels = paste0(CBSA2010_trichot, ", n = ", n_admissions_in_stat_area))
```
 
 
```{r, include = F}
ggplot(data = admissions_to_service_sites_by_CBSA, aes(x = fct_reorder(SERVICES, percent_of_admittees_in_stat_area_using_service_type), y = percent_of_admittees_in_stat_area_using_service_type)) +
  geom_point() +
  geom_segment(aes(x = SERVICES, 
                   xend = SERVICES, 
                   y = 0, 
                   yend = percent_of_admittees_in_stat_area_using_service_type)) +
  geom_text(aes(label = paste0(percent_of_admittees_in_stat_area_using_service_type, "%")), nudge_y = 22, size = 3) +
  ylim(0,100) + 
  facet_wrap(~ CBSA2010_trichot_labels) +
    labs(
    x = "Service Type",
    y = "Percent",
    title = "Service Utilization at Admission to NE Treatment \nCenters in 2017"
  ) +
  coord_flip() +
  theme_classic()
```

Data validation checks.

```{r}
# How many people have missing values for SUB1, SUB2, SUB3, and CBSA.

table_none_or_missing_rural_obs <- NE_admissions_cleaned %>% 
  filter( SUB1 == "None" | SUB1 == "Missing/unknown/not collected/invalid") %>% 
  filter( SUB2 == "None" | SUB2 == "Missing/unknown/not collected/invalid") %>% 
  filter(SUB3 == "None" |  SUB3 == "Missing/unknown/not collected/invalid")


table_none_or_missing_obs <- NE_admissions_cleaned %>% 
  filter( SUB1 == "None" | SUB1 == "Missing/unknown/not collected/invalid") %>% 
  filter( SUB2 == "None" | SUB2 == "Missing/unknown/not collected/invalid") %>% 
  filter(SUB3 == "None" |  SUB3 == "Missing/unknown/not collected/invalid") 



write_csv(table_none_or_missing_obs, "table_none_or_missing_obs.csv")



#### Ensuring unique CASEID's 
NE_admissions_cleaned %>% count(CASEID) %>% filter(n > 1)

#I want to see percent of admissions that list 3 substances by the CBSA regions. Try to get a sense of how common it was to hit the max of reporting options

missing_by_CBSA <- table_none_or_missing_obs$CBSA2010_trichot %>% table() 
metro_missing <- missing_by_CBSA[1] 
micro_missing <- missing_by_CBSA[2] 
outside_missing <- missing_by_CBSA[3]


metro_admissions <- admissions_per_CBSA[1, 2] 
micro_admissions <- admissions_per_CBSA[2, 2]
outside_admissions <- admissions_per_CBSA[3, 2]

percent_metro_missing <- ((metro_missing / metro_admissions) * 100 ) %>% round(2)
percent_micro_missing <- ((micro_missing / micro_admissions) * 100 ) %>% round(2) 
percent_outside_missing <- ((outside_missing / outside_admissions) * 100 ) %>% round(2)

```


